The complex network of trophic interactions in a subAntarctic oceanic Marine Protected

### 2023-05-08 MS_ToSubmit/MS_Burdwood_foodweb.Rmd
2nd version w/ coauthors suggs

address:
  - code: 1
    address: Centro Austral de Investigaciones Cientificas (CADIC-CONICET), Argentina
  - code: 2
    address: Instituto Antartico Argentino (IAA), Argentina
  - code: 3
    address: Instituto de Ciencias Polares, Ambiente y Recursos Naturales (ICPA), Universidad de Tierra del Fuego (UNTDF), Argentina
  - code: 4
    address: Instituto de Investigaciones Marinas y Costeras (IIMYC), Universidad Nacional de Mar del Plata-CONICET, Argentina
  - code: 5
    address: Wildlife Conservation Society, Argentina
  - code: 6
    address: Instituto de Biodiversidad y Biología Experimental y Aplicada (IBBEA), Universidad de Buenos Aires-CONICET, Argentina
  - code: 7
    address: Museo Argentino de Ciencias Naturales “Bernardino Rivadavia”, Argentina
  - code: 8
    address: Instituto de Diversidad y Ecología Animal (IDEA-CONICET)
  - code: 9
    address: Facultad de Ciencias Exactas, Físicas y Naturales, Universidad Nacional de Córdoba (UNC), Argentina

abstract: |
  Globally, the total area of the world ocean designated under marine protection has increased in the recent decades. The majority of these Marine Protected Areas (MPAs) focus on the presence of particularly vulnerable, keystone, or charismatic species, large numbers of endemic species, and/or high biodiversity across taxonomic levels. In the sub-Antarctic region, the level of ocean protection is mainly associated to oceanic islands, except for the MPAs Namuncurá - Banco Burdwood I and II (MPA N-BB, 53º–55ºS and 56º–62ºW, respectively), which are associated to a submarine plateau and its southern adjacent deep slope region, respectively. Here, we present the first detailed analysis of the network of predator-prey interactions (food web) for the MPA N-BB ecosystem. We applied a network approach to characterize the food web in terms of complexity and structure, and identifying the species' role in such a framework. In terms of complexity, the MPA N-BB food web consisted of 1778 predator-prey interactions and 379 species, with a link density of 4.69 and a connectance of 0.01. In terms of structure, almost half of the consumers were omnivores (0.48), and the network displayed a small-world pattern. These findings suggest that the ecosystem might be vulnerable to external perturbations targeting highly connected species, although structural properties might provide resilience and resistance, resulting in a rearranged structure that preserves its original functions. Furthermore, we identified several species as important in terms of different aspects of trophic structure and functioning, and response to perturbations. We suggest that generalist species, mainly fishes, play a crucial role in the ecosystem's bentho-pelagic coupling process and that other species, besides the longtail southern cod *Patagonotothen ramsayi* and the Fueguian sprat *Sprattus fuegensis*, should be considered as relevant energy transfers for the ecosystem. Finally, we argue that the diversity of species, including both the benthic and pelagic habitats, is responsible for seccuring the connectivity within the food web against perturbations, therefore contributing to the structure and stability of the ecosystem.

@@

High-latitude marine ecosystems, such as the MPA N-BB, are considered complex systems in terms of biodiversity and ecological interactions [@Day2013; @Kortsch2019; @Cordone2020; @Trathan2021]. Although there is a robust knowledge about the complexity considering the richness of the benthic and plankton communities in the MPA N-BB ecosystem [@Schejter2016; @Guinder2020; @Schejter2020; @APN2022], a better understanding of species interactions' complexity and structure is needed.
This aspect can be tackled by analyzing one of the most frequent relationships between species: the predator-prey interaction [@Bascompte2009]. The sum of predator-prey or trophic interactions of a particular region is referred to as a food web, representing the roadmap for matter and energy flow in an ecosystem.
In recent years, network approaches have been successfully applied to study complex high-latitude marine ecosystems, improving our knowledge on structure, functioning, and response to environmental/anthropogenic changes [@Kortsch2015; @Cordone2018; @Funes2022; @Marina2023]. Among anthropogenic threats, it is worth mentioning that contaminants like mercury and microplastics have been recently reported as important threats to the MPA N-BB region [@Cossi2021; @DiMauro2022; @Fioramonti2022].

In the present work, we present the first detailed analysis of the network of predator-prey interactions, hereafter food web, for the MPA N-BB ecosystem. For this, we applied a network approach to a highly resolved food web. The objective was twofold: characterize the food web in terms of complexity and structure, and identify the species' role in such a network framework.

@@

We analyzed the MPA N-BB network of trophic interactions, or food web, at two levels: A) network, considering species and interactions of the whole network; and B) species, considering interactions and species related to a particular species (Table 1).

The network-level analysis aims to characterize the food web in terms of complexity and structure. For this, we calculated several network properties commonly used to describe empirical food webs [@Pascual2005]: (1) number of species S; (2) number of interactions or links L; (3) link density L/S; (4) connectance L/S^2; (5) omnivory Omn; and (6) small-world pattern. In order to explore the small-world phenomenon, we analyzed the  characteristic path length (CPL) and the clustering coefficient (CC). The CPL is the average shortest path length between all pairs of nodes [@Watts1998]. Here, CPL was calculated as the average number of nodes in the shortest path $CPL_{Min} (i,j)$ between all pairs of nodes $S(i,j)$ in a network averaged over $S(S-1)/2$ nodes:

@@

Also, we estimated the (7) degree distributions for the food web, prey and predators, and each functional group (e.g., Amphipoda, Ascidiacea, Bivalvia, fish, marine mammals, sea birds, among others). The prey and predator distributions indicate the frequency of prey among predators, and viceversa; the functional group's degree shows the distribution of interactions within groups.

@@

In terms of complexity, the MPA Namuncurá - Banco Burdwood food web (N-BB I and II) consisted of 1796 predator-prey interactions and 380 species, where 93% of them were defined at the species taxonomical level (Figure 2, Table S2). The food web presented a link density (e.g., the average number of interactions per species) of 4.73, and a connectance of 0.01. Almost half of the consumers were omnivores (0.48), feeding on sources at different trophic levels. The food web displayed a small-world pattern, meaning that the path length was lower and the clustering coefficient higher than the random networks (Table 2).

@@

```{r figure3, echo=FALSE, fig.width = 6.5, fig.height = 9, fig.align='center', fig.cap="Degree distributions for the (A) food web, for (B) prey among predators, (C) predators among prey, and (D) for each functional group. Groups are vertically ordered by increasing trophic level; groups with less than 3 species were not plotted (e.g., pelagic fish)."}
# Degree distributions: network, predator, prey & group
# above row (network, prey, predator), below row (group)

@@

Most species in the food web were consumers, 339 out of 379; the rest were primary producers, such as diatoms (phytoplankton), and non-living food sources like detritus and necromass.

We found different relationships between the species trophic level (TL) and the rest of the analysed species-level properties (Figure 4A-D). The most evident significant relationship was with trophic similarity, i.e. the higher the species' TL, the lower the trophic similarity or the higher the uniqueness in terms of trophic role (Figure 4C). Here it is noteworthy to highlight those high-trophic level species (TL > 3.1) with low values of trophic similarity: *Bathyraja macloviana* and *Squalus acanthias* (Chondrichthyans), *Diplopteraster clarki* and *Pteraster sp* (echinoderms), *Phalacrocorax atriceps* and *Eudyptes chrysocome* (sea birds), and *Lagenorhynchus cruciger* (marine mammals) (Table S3).

We also found a significant negative relationship between TL and closeness, however less evident, meaning that low-TL species are relatively closer to any other species in the food web (Figure 4B). Species of genera *Calanus* and *Euphausia*, and species of Brachiopoda, all with TL < 3, registered the highest closeness values (Table S3).

Notably, species of mid-TLs (3-4) showed the highest values of betweenness, meaning that those species participated in the highest number of shortest paths between species (Figure 4A). The following are the species with the highest values: *Patagonotothen ramsayi*, *Dissostichus eleginoides*, *Salilota australis* (fishes), *Doryteuthis gahi* (Cephalopoda), and *Patagonotothen guntheri* (Notothenioid fish) (Table S3).

Considering the topological role, 'module specialist' species were the most frequent and presented a wide TL range (1 - 4.77); 'module hub' was constrained to mid-TL species (2.48 - 3.92); 'module connector' from low to mid-TLs (1 - 3.86); and 'network connector', the least frequent, had all of its species in TL = 1, except for one with TL = 3.47 (Figure 4D, see Figure S2 for species' topological roles in a food web graph framework). Here it is important to highlight the two latter topological roles because they are responsible for linking modules and maintaining the connectivity of the food web: 40 species (5 network connectors + 35 module connectors) from 20 different functional groups with a TL range = 1 - 3.86. The following were the five species with a network connector role: Detritus, Diatoms (benthic, centric and pennate), and *Patagonotothen ramsayi*. The 35 species with a module connector role represent these functional groups: Amphipoda, Bivalvia, Brachiopoda, Bryozoa, Cnidaria, Cumacea, Decapoda, detritus, Diatoms, Dinoflagellates, Echinodermata, fish (bentho-pelagic and demersal Osteichthyes, and Chondrychthyes), Foraminifera, Polychaeta, Porifera, Pycnogonida, sea birds, and zooplankton (see Supplementary Material Table S3 for the identity of the species).

@@

```{r figure4, echo=FALSE, fig.width = 6.5, fig.height = 6.5, fig.align='center', fig.cap="Species-level properties by trophic level: (A) betweenness, (B) closeness, (C) trophic similarity, and (D) topological role. Each point represents a species. Linear regressions for betweenness ($y = x - , R^2 = , p-value < $), closeness ($y = x - , R^2 = , p-value = $) and trophic similarity ($y = x - , R^2 = , p-value = $). Note that for panels A, B and C only species with TLs equal or greater than 2 were considered."}
# Betweenness, Closeness, Trophic similarity & Topological role by Trophic level

@@

It is suggested that the small-world pattern, i.e., a network with short path length and high clustering coefficient, is not frequent in complex marine food webs, mainly due to a low clustering coefficient compared to random networks [@Dunne2002b; @Marina2018a]. However, the food web of the MPA N-BB does display a small-world pattern. Consequences of this could be of great importance in recognizing evolutionary paths and the vulnerability to perturbations [@Montoya2002]. On the one hand, a short path length implies a rapid spread of an impact (e.g., contaminant, population fluctuation, local extinction) throughout the network and, at the same time, more potentially adaptive dynamics in the face of external perturbations [@Montoya2002; @Williams2002]. On the other hand, a high clustering coefficient indicates the formation of subnetworks composed only by the neighbours of particular species. This translates into a greater resistance of the network due to the confinement of perturbations mainly within subnetworks and not spreading between them [@Kortsch2019; @Heer2020]. Overall, a small-world topology provides ecological networks with greater resilience and resistance [@Bornatowski2017; @Dormann2017].

@@

We have identified that most of the consumers in the food web of the MPA N-BB either has a narrow diet or are specialists, while few present a broad diet. The most evident generalist species are *Patagonotothen guntheri* [@CovattiAle2022], *P. ramsayi* [@Fischer2022], *Dissostichus eleginoides* [@Troccoli2020], *Bathyraja brachyurops* [@Belleggia2008a], and *B. griseocauda* [@Bellegia2014], with more than 25 potential prey.
Since these species present mid-trophic positions in the food web (3-4), acting as predator and prey, they might be important links between lower and higher trophic levels. This result is in agreement with the sole analysis, using stable isotopes, that exists so far for the trophic structure of the MPA N-BB [@Riccialdelli2020], and resembles other high-latitude marine systems of the Southwest Atlantic and Antarctic regions [@Arkhipkin2013; @Marina2018]. The importance of these particular generalist species also arises since they feed on the benthic and pelagic habitats [@Troccoli2020; @CovattiAle2022; @Fischer2022], linking these realms and contributing to the vertical carbon flow.

On the other hand, a low number of prey are consumed by many predators in the food web of the MPA N-BB. This suggests that there are dominant food sources from which most consumers depend on and from where the ecosystem energy is being transferred to the upper trophic levels. The most demanded source we identified in this study (i.e. detritus) supports the abundant benthic community of filter-feeders [@Schejter2016]. Furthermore, we found that the second and third-most consumed prey were diatoms and species of *Euphausia*, respectively, which are essential sources for the diverse zooplankton community [@Spinelli2020], mid-TL consumers like the Fuegian sprat *Sprattus fuegensis* [@Padovani2021] and *Patagonotothen ramsayi* [@Fischer2022], and top predators [@RayaRey2022; @Riccialdelli2022].

@@

Ecosystems with a pronounced 'wasp-waist' structure are suggested to present a high trophic redundancy, since many species would show similar trophic habits [@Cury2000]. The significant negative relationship between trophic similarity and trophic level enhances the hypothesis of functional similarity at low and mid-TL species compared to higher TL species for the MPA N-BB food web [@Riccialdelli2020]. At the same time, our results highlight the uniqueness in terms of the trophic role of high-TL predators. Here, not only the expected pelagic animals such as marine mammals and sea birds arise as relevant, but also benthic vertebrate (chondrichthyans *Bathyraja macloviana* and *Squalus acanthias*) and invertebrate species (echinoderms *Diplopteraster clarki* and *Pteraster sp*) are noteworthy. The role that such species play in the MPA N-BB ecosystem is unique and perturbations on them might result in unprecedented changes at the trophic structure and functioning level.

@@

The food web studied in the present work might be more representative of the MPA Namuncurá - Banco Burdwood I ecosystem, which comprises the shallow submarine plateau called Burdwood Bank. This is related to the sampling effort that was conducted during the research cruises in the former MPA compared to the MPA N-BB II (i.e. deep flanks to the south). As a consequence most of the data we used to build the network come from studies performed in the MPA N-BB I. Despite this fact, we decided to build the food web considering both MPAs due to the tight oceanographic and ecological connection that exists among them [references in @APN2022].

@@

The MPA N-BB I presents complex oceanographic conditions that generate an internal spatial heterogeneity, mainly along its longitudinal axis [@Matano2019]. Up to date this heterogeneity has been reflected in phytoplankton and zooplankton communities [@Bertola2018; @GarciaAlonso2020; @Spinelli2020], and in fish assemblages [@Delpiani2020]. Moreover, seasonal variations also occur in some physical and biological aspects of the MPA N-BB I [@GarciaAlonso2018; @Matano2019]. Although we were aware of it, we decided to characterize a food web representing the whole MPA N-BB I year round since this is the first study of its type in the area.

These are the perspectives that we suggest should be considered to improve our knowledge regarding the structure, functioning and stability of the MPA N-BB:

Incorporate habitat heterogeneity, since the benthic community in the BB and deep flanks differs [; @Schejter2021]. This heterogeneity might lead to distinct food web properties, hence differences in ecosystem functioning and response to anthropogenic and environmental perturbations [@Kortsch2019; @Cordone2020].
El caso de las ascidias (Tatián)

Incorporate species traits, like body size and mass, which are known to be important drivers in predator-prey interactions. @LaMesa2016

Simulate perturbations on food web and target species: mercury transfer, microplastic pollution, fisheries.
@Cossi2021; @DiMauro2022; @Fioramonti2022

Estimation of interaction strength considering species and interaction traits (i.e. body size, body mass, interaction dimensionality) to gain insights into food web stability and response to anthropogenic and environmental perturbations.
@Nilsson2016; @Cirtwill2018; @Kortsch2021

>>>>

address:
  - code: 1
    address: Centro Austral de Investigaciones Cientificas (CADIC-CONICET), Argentina
  - code: 2
    address: Instituto Antartico Argentino (IAA), Argentina
  - code: 3
    address: Instituto de Ciencias Polares, Ambiente y Recursos Naturales (ICPA), Universidad de Tierra del Fuego (UNTDF), Argentina
  - code: 4
    address: Instituto de Investigaciones Marinas y Costeras (IIMYC), Universidad Nacional de Mar del Plata-CONICET, Argentina
  - code: 5
    address: Wildlife Conservation Society, Argentina
  - code: 6
    address: Instituto de Biodiversidad y Biologia Experimental y Aplicada (IBBEA), Universidad de Buenos Aires-CONICET, Argentina
  - code: 7
    address: Museo Argentino de Ciencias Naturales “Bernardino Rivadavia”, Argentina
  - code: 8
    address: Instituto de Diversidad y Ecologia Animal (IDEA-CONICET)
  - code: 9
    address: Facultad de Ciencias Exactas, Fisicas y Naturales, Universidad Nacional de Cordoba (UNC), Argentina

abstract: |
  Globally, the total area of the world ocean designated under marine protection has increased in the recent decades. The majority of these Marine Protected Areas (MPAs) focus on the presence of particularly vulnerable, keystone, or charismatic species, large numbers of endemic species, and/or high biodiversity across taxonomic levels. In the sub-Antarctic region, the level of ocean protection is mainly associated to oceanic islands, except for the MPAs Namuncurá - Banco Burdwood I and II (MPA N-BB, 53º–55ºS and 56º–62ºW, respectively), which are associated to a submarine plateau and its southern adjacent deep slope region, respectively. Here, we present the first detailed analysis of the network of predator-prey interactions (food web) for the MPA N-BB ecosystem. We applied a network approach to characterise the food web in terms of complexity and structure, and identifying the species' role in such a framework. In terms of complexity, the MPA N-BB food web consisted of 1780 predator-prey interactions and 375 species, with a link density of 4.75 and a connectance of 0.01. In terms of structure, almost half of the consumers were omnivores (0.48), and the network displayed a small-world pattern. These findings suggest that the ecosystem might be vulnerable to external perturbations targeting highly connected species, although structural properties might provide resilience and resistance, resulting in a rearranged structure that preserves its original functions. Furthermore, we identified several species as important in terms of different aspects of trophic structure and functioning, and response to perturbations. We suggest that generalist species, mainly fishes, play a crucial role in the ecosystem's bentho-pelagic coupling process and that other species, besides the longtail southern cod *Patagonotothen ramsayi* and the Fueguian sprat *Sprattus fuegensis*, should be considered as relevant energy transfers for the ecosystem. Finally, we argue that the diversity of species, including both the benthic and pelagic habitats, is responsible for seccuring the connectivity within the food web against perturbations, therefore contributing to the structure and stability of the ecosystem.

@@

High-latitude marine ecosystems, such as the MPA N-BB, are considered complex systems in terms of biodiversity and ecological interactions [@Day2013; @Kortsch2019; @Cordone2020; @Trathan2021]. Although there is a robust knowledge about the complexity considering the richness of the benthic and plankton communities in the MPA N-BB ecosystem [@Schejter2016; @Guinder2020; @Schejter2020; @APN2022], a better understanding of species interactions' complexity and structure is needed.
This aspect can be tackled by analysing one of the most frequent relationships between species: the predator-prey interaction [@Bascompte2009]. The sum of predator-prey or trophic interactions of a particular region is referred to as a food web, representing the roadmap for matter and energy flow in an ecosystem.
In recent years, network approaches have been successfully applied to study complex high-latitude marine ecosystems, improving our knowledge on structure, functioning, and response to environmental/anthropogenic changes [@Kortsch2015; @Cordone2018; @Funes2022; @Marina2023]. Among anthropogenic threats, it is worth mentioning that contaminants like mercury and microplastics have been recently reported as important threats to the MPA N-BB region [@Cossi2021; @DiMauro2022; @Fioramonti2022].

In the present work, we present the first detailed analysis of the network of predator-prey interactions, hereafter food web, for the MPA N-BB ecosystem. For this, we applied a network approach to a highly resolved food web. The objective was twofold: characterise the food web in terms of complexity and structure, and identify the species' role in such a network framework.

@@

We analysed the MPA N-BB network of trophic interactions, or food web, at two levels: A) network, considering species and interactions of the whole network; and B) species, considering interactions and species related to a particular species (Table 1).

The network-level analysis aims to characterise the food web in terms of complexity and structure. For this, we calculated several network properties commonly used to describe empirical food webs [@Pascual2005]: (1) number of species S; (2) number of interactions or links L; (3) link density L/S; (4) connectance L/S^2; (5) omnivory Omn; and (6) small-world pattern. In order to explore the small-world phenomenon, we analysed the  characteristic path length (CPL) and the clustering coefficient (CC). The CPL is the average shortest path length between all pairs of nodes [@Watts1998]. Here, CPL was calculated as the average number of nodes in the shortest path $CPL_{Min} (i,j)$ between all pairs of nodes $S(i,j)$ in a network averaged over $S(S-1)/2$ nodes:

@@

Also, we estimated the (7) degree distributions for the food web, prey and predators, and each functional group (e.g., Amphipoda, Ascidiacea, Bivalvia, fish, marine mammals, seabirds, among others). The prey and predator distributions indicate the frequency of prey among predators, and viceversa; the functional group's degree shows the distribution of interactions within groups.

@@

In terms of complexity, the MPA Namuncurá - Banco Burdwood food web (N-BB I and II) consisted of 1780 predator-prey interactions and 375 species, where 93% of them were defined at the species taxonomical level (Figure 2, Table S2). The food web presented a link density (e.g., the average number of interactions per species) of 4.75, and a connectance of 0.01. Almost half of the consumers were omnivores (0.48), feeding on sources at different trophic levels. The food web displayed a small-world pattern, meaning that the path length was lower and the clustering coefficient higher than the random networks (Table 2).

@@

```{r figure3, echo=FALSE, fig.width = 6.5, fig.height = 9, fig.align='center', fig.cap="Degree distributions for the (A) food web, for (B) prey among predators, (C) predators among prey, and (D) for each functional group. Groups are vertically ordered by increasing trophic level (following coloration of figure 2); groups with less than 3 species were not plotted (e.g., pelagic fish). All functional groups and the species that comprise them are shown in Supplementary Material (Table S3)."}
# Degree distributions: network, predator, prey & group
# above row (network, prey, predator), below row (group)

@@

We found different relationships between the species trophic level (TL) and the rest of the analysed species-level properties (Figure 4A-D). The most evident significant relationship was with trophic similarity, i.e. the higher the species' TL, the lower the trophic similarity or the higher the uniqueness in terms of trophic role (Figure 4C). Here it is noteworthy to highlight those high-trophic level species (TL > 3.1) with low values of trophic similarity: *Bathyraja macloviana* and *Squalus acanthias* (Chondrichthyans), *Diplopteraster clarki* and *Pteraster sp* (echinoderms), *Phalacrocorax atriceps* and *Eudyptes chrysocome* (seabirds), and *Lagenorhynchus cruciger* (marine mammals) (Table S3).

We also found a significant negative relationship between TL and closeness, however less evident, meaning that low-TL species are relatively closer to any other species in the food web (Figure 4B). Detritus, species of genera *Calanus* and *Euphausia*, and Foraminifera, all with TL < 3, registered the highest closeness values (Table S3).

Notably, species of mid-TLs (3-4.2) showed the highest values of betweenness, meaning that those species participated in the highest number of shortest paths between species (Figure 4A). The following are the species with the highest values: *Patagonotothen ramsayi*, *Salilota australis*, *Dissostichus eleginoides* (fishes), *Doryteuthis gahi* (Cephalopoda), and *Patagonotothen guntheri* (Notothenioid fish) (Table S3).

Considering the topological role, 'module specialist' species were the most frequent and presented a wide TL range (1 - 4.7), as well as 'module hub' species (TL = 1 - 4); 'module connector' was constrained to mid-TLs (2 - 3.86); and 'network connector', the least frequent, had all of its species in low-TLs (1 - 2), except for one with TL = 3.48 (Figure 4D, see Figure S2 for species' topological roles in a food web graph framework). Here it is important to highlight the two latter topological roles because they are responsible for linking modules and maintaining the connectivity of the food web: 41 species (4 network connectors + 37 module connectors) from 19 different functional groups with a TL range = 1 - 3.86. The following were the four species with a network connector role: detritus, benthic diatoms, Foraminifera and *Patagonotothen ramsayi*. The 37 species with a module connector role represented these functional groups: Amphipoda, Bivalvia, Brachiopoda, Bryozoa, Hydrozoa (as 'Cnidaria_benthic'), Copepoda, Cumacea, Decapoda, diatoms, Echinodermata, fish (bentho-pelagic and demersal Osteichthyes, and Chondrychthyes), Polychaeta, Porifera, Pycnogonida (as 'Benthos_Misc') and zooplankton (see Supplementary Material Table S3 for the identity of the species).

@@

```{r figure4, echo=FALSE, fig.width = 6.5, fig.height = 6.5, fig.align='center', fig.cap="Species-level properties by trophic level: (A) betweenness, (B) closeness, (C) trophic similarity, and (D) topological role. Each point represents a species. Linear regressions for betweenness ($y = 74.96x - 121.71, R^2 = 0.05, p-value < 0.01$), closeness ($y = 1.01e-05x - 9.48e-4, R^2 = 0.003, p-value = 0.14$) and trophic similarity ($y = -0.02x + 0.11 , R^2 = 0.08, p-value < 0.01$). Note that for panels A, B and C only species with TLs equal or greater than 2 were considered."}
# Betweenness, Closeness, Trophic similarity & Topological role by Trophic level

@@

It is suggested that the small-world pattern, i.e., a network with short path length and high clustering coefficient, is not frequent in complex marine food webs, mainly due to a low clustering coefficient compared to random networks [@Dunne2002b; @Marina2018a]. However, the food web of the MPA N-BB does display a small-world pattern. Consequences of this could be of great importance in recognizing evolutionary paths and the vulnerability to perturbations [@Montoya2002]. On the one hand, a short path length implies a rapid spread of an impact (e.g., contaminant, population fluctuation, local extinction) throughout the network but, at the same time, more potentially adaptive dynamics in the face of external perturbations [@Montoya2002; @Williams2002]. On the other hand, a high clustering coefficient indicates the formation of subnetworks composed only by the neighbours of particular species. This translates into a greater resistance of the network due to the confinement of perturbations mainly within subnetworks and not spreading between them [@Kortsch2019; @Heer2020]. Overall, a small-world topology provides ecological networks with greater resilience and resistance [@Bornatowski2017; @Dormann2017].

@@

We have identified that most of the consumers in the food web of the MPA N-BB either have a narrow diet or are specialists, while few present a broad diet. The most evident generalist species are *Patagonotothen guntheri* [@CovattiAle2022], *P. ramsayi* [@Fischer2022], *Dissostichus eleginoides* [@Troccoli2020], *Bathyraja brachyurops* [@Belleggia2008a], and *B. griseocauda* [@Bellegia2014], with more than 25 potential prey.
Since these species present mid-trophic positions in the food web (3-4), acting as predator and prey, they might be important links between lower and higher trophic levels. This result is in agreement with the sole analysis, using stable isotopes, that exists so far for the trophic structure of the MPA N-BB [@Riccialdelli2020], and resembles other high-latitude marine systems of the Southwest Atlantic and Antarctic regions [@Arkhipkin2013; @Marina2018]. The importance of these particular generalist species also arises since they feed on the benthic and pelagic habitats [@Troccoli2020; @CovattiAle2022; @Fischer2022], linking these realms and contributing to the vertical carbon flow.

On the other hand, a low number of prey are consumed by many predators in the food web of the MPA N-BB. This suggests that there are dominant food sources from which most consumers depend on and from where the ecosystem energy is being transferred to the upper trophic levels. The most demanded source we identified in this study (i.e. detritus) supports the abundant benthic community of filter-feeders [@Schejter2016]. Furthermore, we found that the second and third-most consumed prey were diatoms and species of *Euphausia*, respectively, which are essential sources for the diverse zooplankton community [@Spinelli2020], mid-TL consumers like the Fuegian sprat *Sprattus fuegensis* [@Padovani2021] and *Patagonotothen ramsayi* [@Fischer2022], and top predators such as the black-browed and grey-headed albatrosses (*Thalassarche melanophris* and *Thalassarche chrysostoma*, respectively) [@Catry2004], and baleen whales (species of the genera Balaenoptera and Eubalaena) [@Valenzuela2018].

@@

Ecosystems with a pronounced 'wasp-waist' structure are suggested to present a high trophic redundancy, since many species would show similar trophic habits [@Cury2000]. The significant negative relationship between trophic similarity and trophic level enhances the hypothesis of functional similarity at low and mid-TL species compared to higher TL species for the MPA N-BB food web [@Riccialdelli2020]. At the same time, our results highlight the uniqueness in terms of the trophic role of high-TL predators. Here, not only the expected pelagic animals such as marine mammals and seabirds arise as relevant, but also benthic vertebrate (chondrichthyans *Bathyraja macloviana* and *Squalus acanthias*) and invertebrate species (echinoderms *Diplopteraster clarki* and *Pteraster sp*) are noteworthy. The role that such species play in the MPA N-BB ecosystem is unique and perturbations on them might result in unprecedented changes at the trophic structure and functioning level.

@@

The food web studied in the present work might be more representative of the MPA Namuncurá - Banco Burdwood I ecosystem, which comprises the shallow submarine plateau called Burdwood Bank. This is related to the sampling effort that was conducted during the research cruises in the former MPA compared to the MPA N-BB II (i.e. deep flanks to the south). As a consequence, most of the data we used to build the network come from studies performed in the MPA N-BB I. Despite this fact, we decided to build the food web considering both MPAs due to the tight oceanographic and ecological connection that exists among them [@APN2022 and references therein].

@@

The MPA N-BB I presents complex oceanographic conditions that generate an internal spatial heterogeneity, mainly along its longitudinal axis [@Matano2019]. Up to date this heterogeneity has been reflected in phytoplankton and zooplankton communities [@Bertola2018; @GarciaAlonso2020; @Spinelli2020], and in fish assemblages [@Delpiani2020]. Moreover, seasonal variations also occur in some physical and biological aspects of the MPA N-BB I [@GarciaAlonso2018; @Matano2019]. Considering both MPAs (N-BB I and II), a seasonal variation in the community composition of marine mammals and seabirds was recorded recently [@Dellabianca2023]. The spatial and seasonal variations in the plankton community might affect the energy and matter flow to higher levels of the food web. This has been recently studied in the vicinity of the MPA N-BB I, in the Beagle Channel, where a differential energy flow pattern of the plankton community has been recognised in two micro-basins of the Channel separated by a sill, each with different physicochemical properties [@Giesecke2021], nutrient concentration [@Latorre2023] as well as in the dominant component of the plankton community [@Bruno2023a; @Presta2023]. Although we were aware of the above, we decided to characterise a food web representing the whole MPA N-BB I year round since this is the first study of its type in the area.

Taking into account the mentioned caveats, and with the aim of improving the knowledge regarding the structure, functioning and stability of the MPA N-BB, we suggest that the future perspectives should: 1) incorporate spatial heterogeneity among MPA N-BB I and II [@Schejter2021], which might lead to distinct food web properties in terms of structure and functioning [@Kortsch2019; @Cordone2020]; 2) include species traits, like body size and mass, since they are known to be important drivers in predator-prey interactions [@Brose2019]; 3) simulate the anthropogenic impacts already present in the MPA N-BB ecosystem (e.g. microplastics, mercury) [@Cossi2021; @DiMauro2022; @Fioramonti2022] as perturbations within the framework of the described complex food web; and 4) estimate the interaction strength of each predator-prey relationship in the food web considering species and interaction traits (i.e. body size, body mass, interaction dimensionality), and species density data [@Nilsson2016; @Pawar2012].
